Thermodynamic Structure of the Solar Corona: Tomographic Reconstructions and MHD Modeling

### 2020-03-12 response_to_review_SOLA-D-20-00015_v1.txt

-----------------------------------------------------------------------------------------
Q2. Figure 3 and 11, bottom panels. I find it difficult to distinguish between the red and magenta colours. Can you replace magenta with a different colour, maybe green?

A2. We agree. Actually this affects all following Figures: 3, 4, 11, and 14. In those, we applied the following changes: Red->Orange and Magenta->Red. These emulate our previous choice of colors in Nuevo et al. (2013) and Lloveras et al. (2017). We like to keep all "up" structures in a "warm" palette. We edited text accordingly all along the manuscript. We believe the colors are now clearly distinguishable in all figures.

-----------------------------------------------------------------------------------------
Q3. Page 14, paragraph starting at line 34. You discuss the latitudinal distribution of short and long loops, and state that the small legs mainly populate the equatorial latitudes, and longer legs at higher latitudes. But this is a skewed result - their spatial positioning was a criteria for their categorisation, so this should not be presented as a scientific result.

A3. Several aspects are involved here, the discussion in the paragraph you indicate strongly relates on that in the final paragraph of Section 2.3. In reading them we find that these texts were confusing as they were, so we edited both. Actually, we do not say that *small* legs populate the equatorial latitudes and *large* ones the mid latitudes. We say that if one classifies the *small-closed* loops in DOWN and UP, the former peak near the equator and the latter peak near mid latitudes (at least for the deep minimum rotation CR-2082). The classification of legs in all four classes in Section 2.3 includes limits for the latitude of their footpoints, to distinguish different layers of the streamer belt (legs of type 0, I and II), as well as to separate field lines of the high latitude CHs (legs of type III). We included two new panels in Figure 3, showing frequency histograms of the latitude of the four types of legs at 1.105 Rsun. Within the range Abs(latitude) < 50 deg, and for the solar minimum target CR-2082, down-legs are peaked near the equator while up-legs peak at mid-latitudes. This is less obvious for CR-2208, although it is clear that down-legs peak in the core of the streamer belt. If one analyses Abs(latitude) > 50 deg it turns out that small-down legs are scarce and there is an extra population of small-up legs (not nearly as much as in the peaks seen at mid lats. Should have we not included the limit of Abs(latitude) < 50 deg for closed-small down and up legs (i.e., legs of type 0 and I), their histograms would still peak at around the same latitudes. This can be seen in Figure 7 of Nuevo et al. 2013), were they analyzed the small closed up and down loops for two rotations of solar minimum, without placing any limits in the latitude of the footpoints. As our main objective here is to be able to analyze progressively outer layers of the streamer belt, we include these limits in the latitude of the footpoints. We edited in full the final paragraph of Section 2.3, as well as the paragraph of Section 3.1 you point out (now split in two, shorter paragraphs). We trust the new graphs and discussions are clear now. Let us know.

@@

-----------------------------------------------------------------------------------------
Q6. The abstract can be shortened somewhat, reducing the number of introductory sentences and with a little less emphasis on the method and future improvements, and more on the science results. In particular: the results of the energy balance analysis and the estimate of energy flux at the coronal base, which are important.

A6. We fully agree. We edited the Abstract, reducing the number of introductory lines from 12 down to 5, and greatly expanding the details in the results, while keeping the full abstract 2 lines shorter than in V1. The presentation of results is better organized now. We first summarize purely DEMT results. For this we were careful to mention which findings were first reported in previous DEMT studies. We still believe these are worth highlighted here as these were now found using the latest implementation of DEMT, in which we implemented here substantial improvements, as discussed in the draft. We close the abstract summarizing how the results of the latest implementation of AWSoM compare with DEMT, highlighting its shortcomings as well.

@@

- Page 26, lines 38-43, between 'Results of the AWSoM….summarized as follows.' can be omitted.

Agreed, done. A little edit was introduced (red text) to start the paragraph.

-----------------------------------------------------------------------------------------
Q10. Minor grammatical improvements, typos, and removal of redundant statements

A10. Thanks very much for this detailed corrections. We basically agreed with all of them and addressed them all. Changes highlighted in red.

### 2020-03-12 lloveras_2020_v2.tex
minimas modificaciones

{As detailed in Section \ref{intro}, rotations CR-2082 and CR-2208 were selected to carry out DEMT reconstructions, based on data taken by the STEREO/EUVI-B and SDO/AIA instruments, respectively.} The EUVI and AIA data were prepared using the latest processing tools and calibration corrections provided by their teams through the SolarSoft package. \edit{In the case of EUVI data, stray-light contamination is removed by deconvolution of the point spread function (PSF), carefully determined for each detector by \citet{shearer_2012}. In the case of AIA data, we have not yet implemented such a procedure as we were not aware of reliable determinations of their PSF, and we also understand that stray light contamination is expected to be not so important for this instrument. A recent study by \citet{saqri_2020} indicates that the effect is noticeable in DEM analysis of coronal holes (CHs). That is also the case for DEM analysis of EUVI images. Nonetheless, as shown by \citet{lloveras_2017}, due to the temporal and spatial binning of the images used in tomography the effect of stray-light removal in DEMT products turns out to be mild, being $\lesssim 10\%$ for density products and negligible for temperature products. In the future, we will explore the effect of stray-light removal in AIA images on DEMT tomography, which we expect to be smaller than for EUVI images.} For this work, we introduced two improvements in the implementation of the DEMT technique, as described next.

@@

{The solution of the tomographic problem involves {inversion of} a very large sparse matrix. Such inversion problems are characterized by spurious {high-fre\-quen\-cy} artifacts in the solution, which can be mitigated through {regularization} techniques \citep{frazin_2000}. In the case of DEMT, all previous efforts used the 2D scheme implemented by \citet{frazin_2009}, using a finite difference matrix operator to approximate angular derivatives in both latitude and longitude. Also new to the present work, is the implementation of an expanded 3D regularization scheme, which adds to the previous scheme a finite difference matrix operator to approximate radial derivatives. In this way, the tomographic inversion problem is performed penalizing nonphysical high-frequency artifacts in all three spatial directions. As a result, tomographic reconstructions behave more smoothly close to the radial boundaries of the computational grid {when compared to previous reconstructions.}}

@@

\edit{In the case of closed-small field lines, the population of down and up legs becomes comparable for CR-2082, so we classify them into the two complementary classes of legs of type 0 (down) and legs of type I (up). In the case of closed-large field lines, legs of type down are $\lesssim 15\%$ of the population for both rotations. In the case of open field lines, legs of type down are $\lesssim 10\%$ of the population for both rotations. Hence, the requirement of being up for legs of type II and III is included to select the vastly dominating population in each case. On the other hand, the inclusion of latitude limits for the footpoints in the classification of legs from type 0 through III is purposely set to study the streamer belt in progressively outer layers, as well as to separate the field lines of the high latitude CHs (legs of type III). In Section \ref{resu}, the results of both the DEMT and AWSoM models in the four classes of legs are statistically analysed.}

@@

{Both target rotations are highly axisymmetric, \textit{i.e.} characterized by a high azymuthal symmetry. Rotation CR-2082} \edit{has} two {small} ARs, both near latitude $+30\mdeg$ and {around longitudes} $50\mdeg$ and $120\mdeg$  {(not identified in the NOAA catalog)}. {Rotation CR-2208} \edit{has} two {ARs, both near latitude} $+5\mdeg$ and {around longitudes $140\mdeg$ and $300\mdeg$ {(NOAA 12722, 12721)}.}

@@

{{Figures \ref{carmaps_demt_2082} and \ref{carmaps_demt_2208} show that} the DEMT reconstruction of the streamer belt is characterized by relatively higher densities and temperatures in comparison to the CHs. They also show that the streamer belt region of CR-2082 was denser and colder than  that of CR-2208.} In the case of CR-2082, which belongs to the deep minimum epoch between {SCs 23 and 24}, the {low latitudes of the streamer belt are characterized by lower electron temperature than in its mid-latitudes.} A similar behavior is seen in CR-2208, belonging to the {end of the declining} phase of {SC 24}, but having a {somewhat less axisymmetric} structure this characteristic is not so obvious. This thermodynamic structure of the streamer has been reported for other solar minimum rotations in previous DEMT works \citep{lloveras_2017,nuevo_2013,vasquez_2010}.

@@

{Throughout the magnetically closed region of both rotations, type 0, I and II legs, associated to  increasingly outer layers of the equatorial streamer belt, exhibit progressively decreasing coronal base density, increasing density scale height, and increasing electron temperature. In both rotations also, type III legs in the CHs are characterized by sub-MK temperatures, and electron density values of order $\approx 1/2$ of those observed for the type 0 and type I legs in the core of the equatorial streamer.} {A comparison of the results between the two rotations shows that, compared to CR-2082, rotation CR-2208 was characterized by {$\approx 10-20\%$} lower values of the electron density at the coronal base, $\approx 5-30\%$ larger values of density scale height, and {$\approx 15-25\%$} larger values of the electron temperature.}

@@

In comparing the DEMT results obtained for the two selected targets, {we highlight} they rely on data provided by two different instruments, {namely EUVI and AIA for CR-2082 and CR-2208, respectively.} To quantify the systematic difference of the DEMT products {due to the different filter sets of} both instruments, \citet{nuevo_2015}, who were the first to apply DEMT to AIA data, {analysed} a single target using both instruments independently. They found that while the density {products are} essentially equal, the temperature \edit{based on AIA data} is systematically 8\% larger than the one based on EUVI data, \textit{i.e.} $\Tm^{\rm(AIA)}/\Tm^{\rm(EUVI)} \approx 1.08$. Considering \edit{this error}, Figure \ref{histos_fulldemt} and Table \ref{tabla_demt} indicate that CR-2208 was {characterized by temperatures $\approx 10-15\%$ larger relative to CR-2082} throughout the streamer belt region. As for the electron density products, CR-2208 was found to be $\approx 10-20\%$ less dense than CR-2082 throughout the streamer belt region. These systematic differences are {around or beyond the uncertainty level in the DEMT products due to systematic sources (radiometric calibration and tomographic regularization), that \citet{lloveras_2017} {estimated to be $\Delta\Tm\approx 10\%$ and $\Delta\sqravgN\approx 5\%$}.}

@@

The sign of the quantity $\phi_c$ depends on that of the conductive flux $F_c$. Equations \ref{Fc} and \ref{phi_c} imply that, by definition, down loops (type 0) and up loops (type I and II) are characterized by $\phi_c<0$ and $\phi_c>0$, respectively, {as verified} in Figure \ref{energia_demt}.

{{Adding the} radiative and conductive terms, the characteristic energy input flux at the coronal base is in the range $\phi_h\approx 0.5-1.5 \times 10^5\,\erg\,\cminvs\,\s^{-1}$, depending on the rotation and the type of {loop, matching the values reported by} \citet{maccormack_2017}. \edit{For} type 0 loops there is a marginal population characterized by the unphysical result $\phi_h<0$. As shown by \citet{maccormack_2017}, this affects only the smallest sized loops of the type 0, and it is likely due \edit{to emission out of the instrumental sensitivity range}. Though accounting for most of the coronal plasma, there surely is additional emission out of the instrumental sensitivity range. As a result,} the positive term $\phi_r$ is most likely underestimated, leading to values $\phi_h<0$ in loops of type 0, {being characterized by $\phi_c<0$.}

@@

{When compared to DEMT results (Figures \ref{carmaps_demt_2082} and \ref{carmaps_demt_2208}), the latitude-longitude maps of the AWSoM model for heights $1.065$ and $1.105\,\mrsun$ capture well the denser and hotter equatorial streamer belt surrounded by the less dense and colder CHs. Furthermore, for both rotations, the temperature maps show the low latitudes of the equatorial streamer belt to be characterized by lower temperatures than its mid-latitudes, as also seen in the DEMT results. The latitude-longitude maps of the AWSoM and DEMT results are shown in the same units and scales, so that a \edit{visual comparison reveals} similar values of electron density and temperature in both models.}

@@

{For both target rotations, the top panels of Figure \ref{rpoint_awsom} show the latitude-longitude location (at heliocentric height $1.105\,\mrsun$) of all traced field line legs for which criterion (i) of Section \ref{trace} is met. That criterion is adapted here, requiring that at least five voxels of the tomographic grid are threaded by the leg. Open legs are indicated in gray color and closed ones in black color. For each leg, the fits to tomographic temperature and density were applied, as given by Equations \ref{Nfit} and \ref{Tfit}. Considering the AWSoM data points and the resulting fits along each leg, the bottom panels of Figure \ref{rpoint_awsom} show the latitude-longitude location of the subset for which also both criteria (ii) and (iii) of Section \ref{trace} are met. Using a three-color code, type I, II and III legs are shown in red, magenta and cyan color, respectively. This figure \edit{can be compaed} with the corresponding Figure \ref{rpoint_demt} for DEMT results. \edit{The AWSoM maps are more densely populated} than those of DEMT. This is due to the 3D MHD model having \edit{having spatially smoother distributions of electron density and temperature than those of DEMT}.}

@@

{For the two target rotations, Table \ref{tabla_comp} summarizes a quantitative comparative analysis between the results of the DEMT and AWSoM models {based on the results shown in Figures \ref{histos_2082} and \ref{histos_2208}}. The DEMT results are expressed as absolute values, while the ASWSoM results are informed as a percentual variation relative to the corresponding result for DEMT. \del{The following major results can be drawn.}}

@@

\edit{Magnetic field lines of type 0, I and II were selected to be associated with increasingly outer layers of the equatorial streamer belt (Figure \ref{rpoint_demt}). These magnetic structures progressively exhibit decreasing coronal base density, increasing density scale height, and increasing electron temperature, as informed in 3D quantitative detail in Figure \ref{histos_fulldemt} and Table \ref{tabla_demt}.} For both rotations we find that down legs populate the low latitudes of the streamer belt, while up legs dominate its mid-latitudes. Also, in the case of CR-2082 the fraction of down legs is significantly larger than for CR-2208. These findings are consistent with previous studies by \citet{huang_2012} and \citet{nuevo_2013}. In the case of the latter, they include in their analysis target CR-2081, which is a rotation almost identical to our target CR-2082. Our results for target CR-2082 compare very well with those of \citet{nuevo_2015} and \citet{lloveras_2017} for target CR-2081. As our study uses the improved version of the DEMT technique, such comparison provided a consistency \edit{check.} \edit{For both rotations, type III field lines in the CHs are characterized by sub-MK temperatures, and electron density values of order $\approx 1/2$ of those observed for the type 0 and type I lines in the core of the equatorial streamer.}

>>>>

{As detailed in Section \ref{intro}, rotations CR-2082 and CR-2208 were selected to carry out DEMT reconstructions, based on data taken by the STEREO/EUVI-B and SDO/AIA instruments, respectively.} The EUVI and AIA data were prepared using the latest processing tools and calibration corrections provided by their teams through the SolarSoft package. \edit{In the case of EUVI data, stray-light contamination is removed by deconvolution of the point spread function (PSF), carefully determined for each detector by \citet{shearer_2012}. In the case of AIA data, we have not yet implemented such a procedure as we were not aware of reliable determinations of their PSF, and we also understand that stray light contamination is expected to be less important for this instrument. A recent study by \citet{saqri_2020} indicates that the effect is noticeable in DEM analysis of coronal holes (CHs). That is also the case for DEM analysis of EUVI images. Nonetheless, as shown by \citet{lloveras_2017}, due to the temporal and spatial binning of the images used in tomography the effect of stray-light removal in DEMT products turns out to be mild, being $\lesssim 10\%$ for density products and negligible for temperature products. In the future, we will explore the effect of stray-light removal in AIA images on DEMT tomography, which we expect to be smaller than for EUVI images.} For this work, we introduced two improvements in the implementation of the DEMT technique, as described next.

@@

{The solution of the tomographic problem involves {inversion of} a very large sparse matrix. Such inversion problems are characterised by spurious {high-fre\-quen\-cy} artifacts in the solution, which can be mitigated through {regularization} techniques \citep{frazin_2000}. In the case of DEMT, all previous efforts used the 2D scheme implemented by \citet{frazin_2009}, using a finite difference matrix operator to approximate angular derivatives in both latitude and longitude. Also new to the present work, is the implementation of an expanded 3D regularization scheme, which adds to the previous scheme a finite difference matrix operator to approximate radial derivatives. In this way, the tomographic inversion problem is performed penalizing nonphysical high-frequency artifacts in all three spatial directions. As a result, tomographic reconstructions behave more smoothly close to the radial boundaries of the computational grid {when compared to previous reconstructions.}}

@@

\edit{In the case of closed-small field lines, the population of down and up legs becomes comparable for CR-2082, so we classify them into the two complementary classes of legs of type 0 (down) and legs of type I (up). In the case of closed-large field lines, down legs are $\lesssim 15\%$ of the population for both rotations. In the case of open field lines, down legs are $\lesssim 10\%$ of the population for both rotations. Hence, the requirement of being up for legs of type II and III is included to select the vastly dominating population in each case. On the other hand, the inclusion of latitude limits for the footpoints in the classification of legs from type 0 through III is purposely set to study the streamer belt in progressively outer layers, as well as to separate the field lines of the high latitude CHs (legs of type III). In Section \ref{resu}, the results of both the DEMT and AWSoM models in the four classes of legs are statistically analysed.}

@@

{Both target rotations are highly axisymmetric, \textit{i.e.} characterised by a high azymuthal symmetry. Rotation CR-2082} \edit{has} two {small} ARs, both near latitude $+30\mdeg$ and {around longitudes} $50\mdeg$ and $120\mdeg$  {(not identified in the NOAA catalog)}. {Rotation CR-2208} \edit{has} two {ARs, both near latitude} $+5\mdeg$ and {around longitudes $140\mdeg$ and $300\mdeg$ {(NOAA 12722, 12721)}.}

@@

{{Figures \ref{carmaps_demt_2082} and \ref{carmaps_demt_2208} show that} the DEMT reconstruction of the streamer belt is characterised by relatively higher densities and temperatures in comparison to the CHs. They also show that the streamer belt region of CR-2082 was denser and colder than  that of CR-2208.} In the case of CR-2082, which belongs to the deep minimum epoch between {SCs 23 and 24}, the {low latitudes of the streamer belt are characterised by lower electron temperature than in its mid-latitudes.} A similar behavior is seen in CR-2208, belonging to the {end of the declining} phase of {SC 24}, but having a {somewhat less axisymmetric} structure this characteristic is not so obvious. This thermodynamic structure of the streamer has been reported for other solar minimum rotations in previous DEMT works \citep{lloveras_2017,nuevo_2013,vasquez_2010}.

@@

{Throughout the magnetically closed region of both rotations, type 0, I and II legs, associated to  increasingly outer layers of the equatorial streamer belt, exhibit progressively decreasing coronal base density, increasing density scale height, and increasing electron temperature. In both rotations also, type III legs in the CHs are characterised by sub-MK temperatures, and electron density values of order $\approx 1/2$ of those observed for the type 0 and type I legs in the core of the equatorial streamer.} {A comparison of the results between the two rotations shows that, compared to CR-2082, rotation CR-2208 was characterised by {$\approx 10-20\%$} lower values of the electron density at the coronal base, $\approx 5-30\%$ larger values of density scale height, and {$\approx 15-25\%$} larger values of the electron temperature.}

@@

In comparing the DEMT results obtained for the two selected targets, {we highlight} they rely on data provided by two different instruments, {namely EUVI and AIA for CR-2082 and CR-2208, respectively.} To quantify the systematic difference of the DEMT products {due to the different filter sets of} both instruments, \citet{nuevo_2015}, who were the first to apply DEMT to AIA data, {analysed} a single target using both instruments independently. They found that while the density {products are} essentially equal, the temperature \edit{based on AIA data} is systematically 8\% larger than the one based on EUVI data, \textit{i.e.} $\Tm^{\rm(AIA)}/\Tm^{\rm(EUVI)} \approx 1.08$. Considering \edit{this error}, Figure \ref{histos_fulldemt} and Table \ref{tabla_demt} indicate that CR-2208 was {characterised by temperatures $\approx 10-15\%$ larger relative to CR-2082} throughout the streamer belt region. As for the electron density products, CR-2208 was found to be $\approx 10-20\%$ less dense than CR-2082 throughout the streamer belt region. These systematic differences are {around or beyond the uncertainty level in the DEMT products due to systematic sources (radiometric calibration and tomographic regularization), that \citet{lloveras_2017} {estimated to be $\Delta\Tm\approx 10\%$ and $\Delta\sqravgN\approx 5\%$}.}

@@

The sign of the quantity $\phi_c$ depends on that of the conductive flux $F_c$. Equations \ref{Fc} and \ref{phi_c} imply that, by definition, down loops (type 0) and up loops (type I and II) are characterised by $\phi_c<0$ and $\phi_c>0$, respectively, {as verified} in Figure \ref{energia_demt}.

{{Adding the} radiative and conductive terms, the characteristic energy input flux at the coronal base is in the range $\phi_h\approx 0.5-1.5 \times 10^5\,\erg\,\cminvs\,\s^{-1}$, depending on the rotation and the type of {loop, matching the values reported by} \citet{maccormack_2017}. \edit{For} type 0 loops there is a marginal population characterised by the unphysical result $\phi_h<0$. As shown by \citet{maccormack_2017}, this affects only the smallest sized loops of the type 0, and it is likely due \edit{to emission out of the instrumental sensitivity range}. Though accounting for most of the coronal plasma, there surely is additional emission out of the instrumental sensitivity range. As a result,} the positive term $\phi_r$ is most likely underestimated, leading to values $\phi_h<0$ in loops of type 0, {being characterised by $\phi_c<0$.}

@@

{When compared to DEMT results (Figures \ref{carmaps_demt_2082} and \ref{carmaps_demt_2208}), the latitude-longitude maps of the AWSoM model for heights $1.065$ and $1.105\,\mrsun$ capture well the denser and hotter equatorial streamer belt surrounded by the less dense and colder CHs. Furthermore, for both rotations, the temperature maps show the low latitudes of the equatorial streamer belt to be characterised by lower temperatures than its mid-latitudes, as also seen in the DEMT results. The latitude-longitude maps of the AWSoM and DEMT results are shown in the same units and scales, so that a \edit{visual comparison reveals} similar values of electron density and temperature in both models.}

@@

{For both target rotations, the top panels of Figure \ref{rpoint_awsom} show the latitude-longitude location (at heliocentric height $1.105\,\mrsun$) of all traced field line legs for which criterion (i) of Section \ref{trace} is met. That criterion is adapted here, requiring that at least five voxels of the tomographic grid are threaded by the leg. Open legs are indicated in gray color and closed ones in black color. For each leg, the fits to tomographic temperature and density were applied, as given by Equations \ref{Nfit} and \ref{Tfit}. Considering the AWSoM data points and the resulting fits along each leg, the bottom panels of Figure \ref{rpoint_awsom} show the latitude-longitude location of the subset for which also both criteria (ii) and (iii) of Section \ref{trace} are met. Using a three-color code, type I, II and III legs are shown in red, magenta and cyan color, respectively. This figure \edit{can be compared} with the corresponding Figure \ref{rpoint_demt} for DEMT results. \edit{The AWSoM maps are more densely populated} than those of DEMT. This is due to the 3D MHD model having \edit{having spatially smoother distributions of electron density and temperature than those of DEMT}.}

@@

{For the two target rotations, Table \ref{tabla_comp} summarizes a quantitative comparative analysis between the results of the DEMT and AWSoM models {based on the results shown in Figures \ref{histos_2082} and \ref{histos_2208}}. The DEMT results are expressed as absolute values, while the ASWSoM results are informed as a percentual variation relative to the corresponding result for DEMT.}

@@

\edit{Magnetic field lines of type 0, I and II were selected to be associated with increasingly outer layers of the equatorial streamer belt (Figure \ref{rpoint_demt}). These magnetic structures progressively exhibit decreasing coronal base density, increasing density scale height, and increasing electron temperature, as informed in 3D quantitative detail in Figure \ref{histos_fulldemt} and Table \ref{tabla_demt}.} For both rotations we find that down legs populate the low latitudes of the streamer belt, while up legs dominate its mid-latitudes. Also, in the case of CR-2082 the fraction of down legs is significantly larger than for CR-2208. These findings are consistent with previous studies by \citet{huang_2012} and \citet{nuevo_2013}. In the case of the latter, they include in their analysis target CR-2081, which is a rotation almost identical to our target CR-2082. Our results for target CR-2082 compare very well with those of \citet{nuevo_2015} and \citet{lloveras_2017} for target CR-2081. As our study uses the improved version of the DEMT technique, such comparison provided a consistency \edit{check.} \edit{For both rotations, type III field lines in the CHs are characterised by sub-MK temperatures, and electron density values of order $\approx 1/2$ of those observed for the type 0 and type I lines in the core of the equatorial streamer.}
